What is essential? -- A pilot survey on views about the requirements metamodel of reqT.org

### 2016-03-14 paper/submitted/paper-refsq2016-final.tex
rename dirs

\textbf{Subject background.} The background questions in the survey regards the role of the subject, as shown in Table \ref{table:background}. The analyzed\footnote{One subject answered NO on all background questions and was therefore excluded.} \input{summary.tex} The response rate was 100\% after a reminder was emailed to one missing subject. %\vspace{-0.6cm}

%\noindent 
\textbf{Frequency analysis.} The degree of ''essentiality'' is characterized as the number of subjects that has responded that they (1) use the concept at least in an informal, non-persistent way, \textit{and} that they (2) use the concept in a similar meaning as in the definition in Appendix A. Fig.~\ref{fig:survey} shows the definitions of the three-level ordinal scales of Questions $Q1_{usage}$  and $Q2_{meaning}$ respectively. Table \ref{table:frequency} shows the results of the frequency counts.  
If an ''essentiality threshold'' is chosen at $N/2$ then only the $9$ concepts from row $n = 7$ and below in Table~\ref{table:frequency} are considered ''non-essential'', hence showing that more than 90\% of the metamodel concepts have a majority of the subjects that use them and agree upon their definitions. Each concept has at least one subject that uses it and agrees with its definition.

The following 19 concepts were reported ''missing'': 
S01: \emph{or}, 
S02: \emph{bug}, \emph{threshold},
S04: \emph{role}, \emph{problem}, \emph{motivates}, \emph{and}, \emph{or}, \emph{pattern}, \emph{submodel},
S06: \emph{plug-in}, \emph{informalism}, 
S07: \emph{full sentence}, 
S09: \emph{satisfaction}, \emph{satisfies}, \emph{customer},
S11: \emph{system-of-interest}, \emph{verification}, \emph{validation}, S13: \emph{context}. 
Thus, the  concept 'or' was the only concept that had consensus among several subjects (S01, S04) as considered ''missing''.

{The anonymised data and analysis scripts (developed using Scala and Apache POI) are available at: \footnotesize\url{https://github.com/bjornregnell/reqT-survey}} 
\vspace{-0.3cm}

\begingroup
\begin{table}[H]
\setlength{\tabcolsep}{4pt} % Default value: 6pt
\renewcommand{\arraystretch}{1.4} % Default value: 1
\centering
\fontsize{8.5}{9}\selectfont
\caption{Background of subjects, $N = 15$. The subjects were given anonymous ids S01--S15.}
\label{table:background}
\vspace{-0.2cm}
\begin{tabular}{p{0.42\textwidth}| p{0.54\textwidth}}
\textit{Background question} & \textit{Subject responding YES}  \\ \hline
\input{background.tex}
 \end{tabular}
\end{table}
\endgroup

\vspace{-3.6em}
\begingroup
\setlength{\tabcolsep}{4pt} % Default value: 6pt
\renewcommand{\arraystretch}{1.35} % Default value: 1
\begin{table}
%\centering
\begin{center}
\fontsize{7}{8}\selectfont
\caption{Frequency analysis, where $n$ is the number of subjects that for the respective concept answered $(Q1_{usage}>= 1)$ \textit{and} $(Q2_{meaning} = 2)$. In total there are 92 concepts (49 entities, 15 relations and 28 attributes). The higher up in the table, the more ''essential''. For $n = 0, 2, 3, 5$ the were no concepts with answers by that number of subjects.}
\label{table:frequency}
\begin{tabular}{l | p{0.33\textwidth} | p{0.30\textwidth} | p{0.25\textwidth}}
\textit{$n$} & \textit{Entities} & \textit{Attributes} & \textit{Relations} \\ \hline
\input{essential.tex}
\end{tabular}
\end{center}

%\vspace{0.1cm}
%\fontsize{9}{10}\selectfont 
%\textbf{Table 3.} $Q3$ \emph{Missing Concepts} according to subjects:
%S01: or, 
%S02: bug, threshold,
%S04: role, problem, motivates, and, or, pattern, submodel,
%S06: plug-in, informalism, 
%S07: full sentence, 
%S09: satisfaction, satisfies, customer,
%S11: system-of-interest, verification, validation, S13: context. 
\end{table}
\endgroup

@@

\textbf{\textit{Entity}} & \textbf{\textit{Definition}}  \\
\input{entity-defs.tex}
 \end{tabular}
\end{table}
\end{minipage}
 \hspace{4em plus 1fill} 
\begin{minipage}{0.46\textwidth} 
 \begin{table}[H]
%\centering
\fontsize{6}{6.5}\selectfont
%\caption{Definitions of attributes and relations.}
\vspace{-1.0cm}
\label{table-entities}
\begin{tabular}{|l p{4.1cm}}

\textbf{\textit{Attribute}} & \textbf{\textit{Definition}}  \\ 
\input{attribute-defs.tex}
 & \\
 \textbf{\textit{Relation}} & \textbf{\textit{Definition}}  \\ 
\input{relation-defs.tex}

>>>>

\textbf{Subject background.} The background questions in the survey regards the role of the subject, as shown in Table \ref{table:background}. The analyzed\footnote{One subject answered NO on all background questions and was therefore excluded.} \input{../analysis/summary.tex} The response rate was 100\% after a reminder was emailed to one missing subject. %\vspace{-0.6cm}

%\noindent 
\textbf{Frequency analysis.} The degree of ''essentiality'' is characterized as the number of subjects that has responded that they (1) use the concept at least in an informal, non-persistent way, \textit{and} that they (2) use the concept in a similar meaning as in the definition in Appendix A. Fig.~\ref{fig:survey} shows the definitions of the three-level ordinal scales of Questions $Q1_{usage}$  and $Q2_{meaning}$ respectively. Table \ref{table:frequency} shows the results of the frequency counts.  
If an ''essentiality threshold'' is chosen at $N/2$ then only the $9$ concepts from row $n = 7$ and below in Table~\ref{table:frequency} are considered ''non-essential'', hence showing that more than 90\% of the metamodel concepts have a majority of the subjects that use them and agree upon their definitions. Each concept has at least one subject that uses it and agrees with its definition. Table~3 shows 19 concepts that subjects reported as ''missing'' ($Q3$), where 'or' was the only concept that had consensus among several subjects as considered ''missing''. {The anonymised data and analysis scripts (developed using Scala and Apache POI) are available at: \footnotesize\url{https://github.com/bjornregnell/reqT-survey}} 
\vspace{-0.3cm}

\begingroup
\begin{table}[H]
\setlength{\tabcolsep}{4pt} % Default value: 6pt
\renewcommand{\arraystretch}{1.4} % Default value: 1
\centering
\fontsize{8.5}{9}\selectfont
\caption{Background of subjects, $N = 15$. The subjects were given anonymous ids S01--S15.}
\label{table:background}
\vspace{-0.2cm}
\begin{tabular}{p{0.42\textwidth}| p{0.54\textwidth}}
\textit{Background question} & \textit{Subject responding YES}  \\ \hline
\input{../analysis/background.tex}
 \end{tabular}
\end{table}
\endgroup

\vspace{-3.2em}
\begingroup
\setlength{\tabcolsep}{4pt} % Default value: 6pt
\renewcommand{\arraystretch}{1.35} % Default value: 1
\begin{table}
%\centering
\begin{center}
\fontsize{7}{8}\selectfont
\caption{Frequency analysis, where $n$ is the number of subjects that for the respective concept answered $(Q1_{usage}>= 1)$ \textit{and} $(Q2_{meaning} = 2)$. In total there are 92 concepts (49 entities, 15 relations and 28 attributes). The higher up in the table, the more ''essential''. For $n = 0, 2, 3, 5$ the were no concepts with answers by that number of subjects.}
\label{table:frequency}
\begin{tabular}{l | p{0.33\textwidth} | p{0.30\textwidth} | p{0.25\textwidth}}
\textit{$n$} & \textit{Entities} & \textit{Attributes} & \textit{Relations} \\ \hline
\input{../analysis/essential.tex}
\end{tabular}
\end{center}

\vspace{0.1cm}
\fontsize{9}{10}\selectfont 
\textbf{Table 3.} $Q3$ \emph{Missing Concepts} according to subjects:
S01: or, 
S02: bug, threshold,
S04: role, problem, motivates, and, or, pattern, submodel,
S06: plug-in, informalism, 
S07: full sentence, 
S09: satisfaction, satisfies, customer,
S11: system-of-interest, verification, validation, S13: context. 
\end{table}
\endgroup

@@

\textbf{\textit{Entity}} & \textbf{\textit{Definition}}  \\
\input{../analysis/entity-defs.tex}
 \end{tabular}
\end{table}
\end{minipage}
 \hspace{4em plus 1fill} 
\begin{minipage}{0.46\textwidth} 
 \begin{table}[H]
%\centering
\fontsize{6}{6.5}\selectfont
%\caption{Definitions of attributes and relations.}
\vspace{-1.0cm}
\label{table-entities}
\begin{tabular}{|l p{4.1cm}}

\textbf{\textit{Attribute}} & \textbf{\textit{Definition}}  \\ 
\input{../analysis/attribute-defs.tex}
 & \\
 \textbf{\textit{Relation}} & \textbf{\textit{Definition}}  \\ 
\input{../analysis/relation-defs.tex}
